Towards Practical First-Order Model Counting

### 2024-04-27 doc/paper.tex
Conclusion (& the first complete draft)

TODO: Prove (lots of notes, handwritten and elsewhere)

@@

TODO: Introduce greedy search and breadth-first search (BFS), together with
\Cranegreedy{} and \Cranebfs{}. The knowledge compilation part of both
\textsc{Crane} and \Cranetwo{} can be executed using either greedy (similar to
\textsc{ForcLift}) or breadth-first search. We use both in our experiments,
denoting them as \Cranegreedy{} and \Cranebfs{}, respectively.

We compare \Cranetwo{} (in both BFS and greedy modes) with \textsc{FastWFOMC}
and \textsc{ForcLift} on two problems previously considered by Dilkas and
Belle~\shortcite{DBLP:conf/kr/DilkasB23}: the function-counting problem from
\cref{example:functions} and the bijection-counting problem described below.
Note that comparing \Cranetwo{} and \textsc{FastWFOMC} on a larger set of
benchmarks is challenging because there is no automated way to translate a
formula in \FO{} or \Ctwo{} into \UFO{} (or even check if such an encoding is
possible). The experiments were run on an AMD~Ryzen~7~5800H processor with
\SI{16}{\gibi\byte} of memory and Arch Linux~6.8.2-arch2-1 operating system.
\textsc{FastWFOMC} was executed using Python~3.8.19 with Python-FLINT~0.5.0.

@@

\section{Conclusion}

TODO

>>>>

@@

We compare \Cranetwo{}\footnote{\textsc{Crane} has two modes that decide how
  compilation rules are applied to formulas: one based on greedy search and one
  that combines greedy and breadth-first search. We include both modes in our
  experiments, referring to them as \Cranegreedy{} and \Cranebfs{},
  respectively.} with \textsc{FastWFOMC} and \textsc{ForcLift} on two problems
previously considered by Dilkas and Belle~\shortcite{DBLP:conf/kr/DilkasB23}:
the function-counting problem from \cref{example:functions} and the
bijection-counting problem described below. Note that comparing \Cranetwo{} and
\textsc{FastWFOMC} on a larger set of benchmarks is challenging because there is
no automated way to translate a formula in \FO{} or \Ctwo{} into \UFO{} (or even
check if such an encoding is possible). The experiments were run on an
AMD~Ryzen~7~5800H processor with \SI{16}{\gibi\byte} of memory and Arch
Linux~6.8.2-arch2-1 operating system. \textsc{FastWFOMC} was executed using
Python~3.8.19 with Python-FLINT~0.5.0.

@@

\section{Conclusion and Future Work}

In this work, we presented:
\begin{enumerate*}[label=(\roman*)]
  \item algorithmic techniques for finding the base cases of recursive functions
  produced by the original \textsc{Crane} algorithm,
  \item an extension to the smoothing procedure of \textsc{ForcLift} and
  \textsc{Crane} that supports base case formulas and the compilation rules
  introduced by Dilkas and Belle~\shortcite{DBLP:conf/kr/DilkasB23}.
  \item an approach for compiling function definitions to C++ programs with
  support for arbitrary-precision arithmetic, and
  \item experimental evidence showing that \Cranetwo{} can scale to domain sizes
  orders of magnitude higher than \textsc{FastWFOMC} while being able to handle
  more formulas than \textsc{ForcLift}.
\end{enumerate*}

Many avenues for future work remain. In particular, a more extensive
experimental study is warranted to better understand how WFOMC algorithms
compare in terms of
\begin{enumerate*}[label=(\roman*)]
  \item what formulas they can handle and
  \item their scalability with respect to domain size.
\end{enumerate*}
The capabilities of \Cranetwo{} could be further characterised in other ways.
For instance, one could prove \emph{completeness} for a fragment of first-order
logic such as \Ctwo{} (via a suitable encoding of counting quantifiers).
Moreover, the efficiency with which a WFOMC algorithm is able to handle a
particular formula can be assessed using \emph{fine-grained complexity}. In the
case of \textsc{Crane} and \Cranetwo{}, one can do so simply by analysing the
equations~\cite{DBLP:conf/kr/DilkasB23}. This way, one can reason about
efficiency in a more implementation-independent manner by making claims about
the maximum degree of the polynomial that characterises any particular solution.
